Interference Transform: Estimating the frequency and phase of low resolution samples

### 2020-08-03 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
progress in the paper

\documentclass{article}
\usepackage{amsfonts} % \mathbb
\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\usepackage{amsmath}
\DeclareMathOperator{\sign}{sign}

@@

\begin{abstract} % 250 words
  % background estimate parameters from the signal itself
  Despite being an elusive concept, only well defined mathematically in artificially generated waves, the envelope of a signal is essential for its complete characterization, being the primary information carrying medium in telecommunications, for example.
  the concept of an envelope is not uniformly defined across the literature and can be understood intuitively as a smooth function of the same variable as the principal wave that modulates the signal, responsible for his outer shape. It is implied in this definition that the envelope modulates the instantaneous amplitude of the underlying signal, being generally non periodic and thus better addressed by specific techniques diverse from those used in the wave it encompasses.
  % motivation
  Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology, speech recognition.
  % objective
  In this paper we propose a framework that uses information about the curvature of a signal to identify its boundary, eliminating the necessity of parameter tuning.
  % methods
  The approach here described draws inspiration from geometric algorithms, most notably the concept of alpha-shapes and concave hulls, to identify the frontier of an arbitrary signal. By defining a pulse as the atomic entity of a signal we greatly reduce the dimensionality and randomness of the problem, enabling to asses estimates about the curvature of a signal to inform in the envelope identification.
  % results

@@

@@

\section{Methodology} % 1000

We start by defining a pulse as a series of consecutive samples in a discrete wave with the same signal. More formally, let $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}_0, i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a subset of the natural numbers. This definition relates closely to the concept of an array or vector in programming languages, and is used in the interest of simplicity. A pulse $ P[j], j \in \mathbb{N}_0 $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall a \le i \le b | a,b \in \mathbb{N}_0 $; that is to say, a cluster of samples between the change of sign. For a continuous function, that would correspond to the pieces between roots, and would be equal to half the cycle of a sinusoid.

It is useful now to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that lie on the smooth, continuous envelope of the signal forming, in this way, a discrete version of the envelope itself. Because we are interested in a smooth function over the whole domain of the wave, we can assume that, in the context of a pulse, this function will be very close to a straight line and, thus, won't touch any point inside the convex hull defined by the pulse. By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the envelope, one can safely assume that $ \max(P[j]) \approx \max(P[j]+1) $ and approximate the convex hull of a pulse with the triangle defined by the points where it crosses the horizontal axis and it's interior point of max amplitude. In practice, this will allow a considerable dimension reduction.

To obtain an estimate of the behaviour of the envelope we observe the evolution of it's instantaneous curvature. For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is defined as $c(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $. As we are dealing with a discrete variable, we may use a finite differences scheme with modifications to account for the fact that, in general, the distance from pulse to pulse is variable.

@@

\section{References}

>>>>

\documentclass{article}
\usepackage{amsfonts} % \mathbb
\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[backend=biber, style=apa]{biblatex}
\usepackage{amsmath}
\DeclareMathOperator{\sign}{sign}
\bibliography{bibli}

@@

\begin{abstract} % 250 words

% estimate parameters from the signal itself
% Dominant frequency as by-product
% we define several entities in the process, introducing a new measure of discrete curve curvature, hoping to consolidate the terminology in the area

  % background  
  Despite being an elusive concept, only mathematically well defined in artificially generated waves, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
  intuitively, the temporal envelope can be understood as a smooth function on the same variable as the principal wave, or temporal fine structure, that modulates the signal, responsible for its outer shape. It is implied in this definition that the envelope modulates the instantaneous amplitude of the underlying signal, being non periodic in general and thus better addressed by specific techniques diverse from those used to analyse the wave it encompasses.
  % motivation
  Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
  % objective
  In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
  % methods
  The approach here described draws inspiration from geometric algorithms, most notably the concept of alpha-shapes and concave hulls, to identify the frontier of an arbitrary signal. By defining a pulse as the atomic entity of a signal we greatly reduce the dimensionality and randomness of the problem.
  % results

@@

% Opener sentence
% Literature review | Brief Context of Prior Research
In many applications the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to Qi et al (\citedate{2017QiRelative}), for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones. That is also the case for English, according to \cite{1995ShannonSpeech}, where even envelopes modulating mostly noise were capable of conveying meaning. Still in the context of speech, the envelope helps to convey emotion and identity \parencite{2018ZhuContributions}.

Also, the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a good description of the envelope would be helpful for a subsequent spectral analysis.

% Restate Your Question as Something Not Known or Fully Understood by Prior Research
% State the Significance of Your Question
% State Your Claim | Objective | Hypothesis
% brief outlook on the structure of the paper

@@

\section{Methodology} % 1000 words

% Essential background information
We start by defining a pulse as a series of consecutive samples in a discrete wave with the same signal. More formally, let $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}_0, i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a subset of the natural numbers. This definition relates closely to the concept of an array or vector in programming languages, and is used in the interest of simplicity. A pulse $ P[j], j \in \mathbb{N}_0 $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall a \le i \le b | a,b \in \mathbb{N}_0 $; that is to say, a cluster of samples between the change of sign. For a continuous function, that would correspond to the pieces between roots, and would be equal to half the cycle of a sinusoid.

As we are seeking a geometric treatment of the subject, it is necessary to make abscissa and ordinate units consistent. this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same time related units.

It is useful now to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that bound the smooth, continuous envelope of the signal. Because we are interested in a smooth function over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a straight line and, thus, won't touch any point inside the convex hull defined by the pulse. By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the envelope, one can assume that $ \max(P[j]) \approx \max(P[j+1]) $ and approximate the convex hull of a pulse with the triangle defined by the points where it crosses the horizontal axis and it's point of max amplitude; this will allow a considerable dimension reduction as this triangular pulse becomes the atomic unit for the rest of the method.

For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $c(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $. This definition provides a welcome link between algebra and geometry, but also poses two difficulties to our particular case that need to be addressed: the fact that our signal isn't even continuous, and thus neither smooth, and that we don't have a solid unit for the amplitude.

@@

\section{References}
\printbibliography[heading=none]
